### 2021-02-26 yfcc100m/paper/vldb2021/eval.tex
Update eval

Given that there are no other open-source systems that provide similar
functionality and interfaces as VDMS (i.e., transactionally dealing with images
and metadata behind a single interface), we have implemented two equivalent
visual data management system as baselines, comprised of a combination of
widely available, off-the-shelf components.
The first baseline uses MySQL Server 5.7 (for storing metadata),
Apache Web Server 2.4.18 (as interface for image access), and
OpenCV 3.3 (to provide pre-processing operations on images).
The other baseline system replaces the MySQL Server 5.7 with a
most advanced open-source relational database, PostgreSQL 9.5 \cite{postgresql}.
We decided to use relational databases in the baseline systems instead of
non-relational databases because of their madurity, and also because it is
the most common tools used for storing and querying metadata
~\cite{Jatana2012, li_2019}:

@@

For all our experiments, we use two servers with Ubuntu 20.04,
one hosting a VDMS server and another hosting the baseline implementations.
Both servers have a dual-socket Intel\textsuperscript{\textregistered}
Xeon\textsuperscript{\textregistered} Platinum 8180 CPU @ 2.50GHz (Skylake),
each CPU with 28 physical cores with hyper-threading enabled,
for a total of 112 logical cores per server.
The server hosting MySQL and PostgreSQL has 256GB of DDR4 DRAM,
while the server hosting VDMS has 194GB of DDR4 DRAM.
We decided to run the VDMS server in the machine with less DRAM to make
sure MySQL and PostgreSQL had no disadvantage, and because previous evaluation
indicated smaller footprint in the case of VDMS when
compared to similar baselines based on MySQL.
To build both MySQL and PostgreSQL on the same machine,
we stored each relational database on separate SSD drives.
Other than the difference in DRAM space and storage needed for two baselines,
the machines are identical.
The client application running the queries and measuring round-trip time
is connected to the server through a 1GB wired link through
a 10GB back-plane switch, same as both servers.
The client application was implemented using Python 3 for both VDMS and the baselines.
Figure~\ref{fig:systems} shows a logical view of the difference between the
interaction of the client application (retrieves metadata and
images) with VDMS (left) and the MySQL/PostgreSQL baselines (right).

@@

% Another important aspect to note is the storage footprint of the databases.
PostgreSQL and VDMS require comparable amount of storage for metadata which
is more than required by MySQL, shown in Figure~\ref{fig:db_sizes}.
In the case of VDMS, this is space used to store information about
each node/connection.
The Graph Database internal to VDMS (called PMGD) was designed for performance,
especially in environments where persistent memory is present.
This design decision comes as a trade-off for storage footprint, which is
noticeable in our results.
VDMS required 37-41\% more storage than MySQL for storing the same amount
of metadata. On the other hand, PostgreSQL required 1-4\% more storage than VDMS.
The storage footprint may become a factor if storage is a limitation, but
it should also be noted that metadata accounts for less than 2\% of the
overall database size even if we have a 41\% increase in metadata size.
For example, the largest database (both metadata and images) we built
(100M) has around 230GB of metadata and 12TB of images.
In systems where persistent memory is a scarce resource,
the increased storage foot print of PMGD may represent a challenge.
On the other hand, persistent memory is expected to be available
in the order of TBs per server, which should fit the
metadata of intensive use-cases\cite{IntelXPoint15}.

@@

\begin{figure}[ht]
% WAS: begin{figure*}[ht]; textwidth
\includegraphics[width=\columnwidth]{figures/plot_conc_q_1tag_geo_resize_mosaic_results_throughput_db_size}
\caption{Concurrency Analysis on \textit{q2} (\textit{1tag\_geo\_resize}) from our use-case
described in the Experimental Setup Section.
We show each database in different figures for readability reasons.
The figures show the throughput (images per second) when retrieving
resized versions of the images, as the number of concurrent clients increases
Hardware concurrency (number of physical cores in each system)
is shown with a blue vertical line (hw-concurrency = 56).}
\label{fig:concurrency_comparison_q2}
\end{figure}

@@

Another thing to notice is that, as the number of concurrent clients increases,
VDMS throughput continues to increase up to 56 threads, which is
the hardware concurrency of the system.
Also, more parallelism after 56 threads does not increase the delivered throughput
for the larger databases (50M and 100M) but there is a slight improvement in the
other databases (1M, 5M, and 10M).
On the other hand, both baselines seem to deliver less aggregated
throughput after 16 threads except for the 100M case.
In this case, PostgreSQL has a performance spike at 32 clients
and the performance for both baselines is less stable when compared
to VDMS.
Note that most of the throughput for the baseline systems are very close to each other.
This is mostly an effect of the log-scale used, which is needed to clearly
depict the difference between VDMS and the baselines.
Here, the baselines are the full architecture described in
Figure~\ref{fig:systems} (right).

@@

\begin{figure}[ht]
% WAS: begin{figure*}[ht]; textwidth
\includegraphics[width=\columnwidth]{figures/plot_conc_dbsize_10M_mosaic_results_throughput}
\caption{Concurrency Analysis for all the queries using the 10M database).
We show queries in different figures for readability reasons.
Each figure shows the throughput (images per second) when retrieving
resized versions of the images as the number of concurrent clients increases.
Hardware concurrency (number of physical cores in each system)
is shown with a blue vertical line (hw-concurrency = 56).}
\label{fig:concurrency_comparison_10M}
\end{figure}

We continue by looking at Figure~\ref{fig:concurrency_comparison_10M}
which shows the images per second delivered by each system
for all queries using the 10M database, as the number of concurrent
clients grows (x-axis). This figure takes a closer look at the
concurrency for each of the six queries.
It is obvious that VDMS consistently outperforms each of the baseline
systems for queries \textit{q1}, \textit{q2}, \textit{q4},
and \textit{q6} as concurrency grows. This figure
shows a similar trend as Figure~\ref{fig:concurrency_comparison_q2}
when it comes to concurrency.  The
VDMS throughput continues to increase up to 56 concurrent clients, which is
the hardware concurrency of the system, for all queries except \textit{q3}.
In this query, the throughput seems to decrease after 32 clients
and both baselines outperform VDMS after 8 concurrent clients.
Prior to 8 clients, VDMS has a slight advantage over PostgreSQL which has
a major performance improvement at 16 concurrent clients. On the other hand,
MySQL maintains its improvement in throughput over PostgreSQL and VDMS over all
concurrent clients. In the case of \textit{q5}, the throughput of the baselines
are less aggregated than those of VDMS up to 16 concurrent clients. When it comes
to low concurrency, the baselines do as good and even better than VDMS with 2 or
4 concurrent clients.  At 16 concurrent client, the performance of the MySQL baseline
begin to stabilizes while the throughput of PostgreSQL begins to degrade. However, as
concurrency increases beyond 16 concurrent clients, the difference
in throughput becomes clear, with VDMS reaching its peak performance at
112 concurrent clients.

@@

For \textit{q1} and \textit{q2}, we can appreciate higher performance being delivered by
VDMS when compared to the baselines, and how this improvement is maintained
as the size of the database increase while
on the other hand, the baseline implementations have a drop in performance
for the 100M database.
For \textit{q3}, we see that VDMS performs best when the database size
is small (1M images), but as the database size increases,
the performance degrades as well.
For this query, MySQL (and in some cases PostgreSQL) outperforms VDMS
for databases larger than 1M.
This also occurs for \textit{q5} but only for databases larger than 10M.
This is entirely attributed to the 2-round process needed for this query,
as we explained before which is more visible in the larger databases.
It is interesting to note that adding filtering by geo-location
(\textit{q5}) decreases the performance as the database sizes scales.
The behaviour is different in the case of \textit{q6},
which also filters by geo-location.
This is attributed to the fact that OR queries involves processing
larger results, and thus does not benefit from the filtering
which is evident when compared to \textit{q4}.

@@

Because of space constraints, we are only showing a subset of our evaluation.

We are making available the all available  of For more details about our comprehensive evaluation.
available
\footnote{https://github.com/luisremis/visual\_storm/tree/master/yfcc100m/python/eval/results}
% TODO. Upload the final version of the plots once we are sure we are not making
% more modifications.

>>>>

Given that there are no other open-source systems that provide similar
functionality and interfaces as VDMS (i.e., transactionally dealing with images
and metadata behind a single interface), we have implemented two equivalent
visual data management system as baselines, comprised of a combination of
widely available, off-the-shelf components.
The first baseline uses MySQL Server 5.7 (for storing metadata),
Apache Web Server 2.4.18 (as interface for image access), and
OpenCV 3.3 (to provide pre-processing operations on images).
The other baseline system replaces the MySQL Server 5.7 with a
most advanced open-source relational database, PostgreSQL 9.5 \cite{postgresql}.
We decided to use relational databases in the baseline systems instead of
non-relational databases because of their maturity, and also because it is
the most common tools used for storing and querying metadata
~\cite{Jatana2012, li_2019}:

@@

For all our experiments, we use two servers with Ubuntu 20.04,
one hosting a VDMS server and another hosting the baseline implementations.
Both servers have a dual-socket Intel\textsuperscript{\textregistered}
Xeon\textsuperscript{\textregistered} Platinum 8180 CPU @ 2.50GHz (Skylake),
each CPU with 28 physical cores with hyper-threading enabled,
for a total of 112 logical cores per server.
The server hosting MySQL and PostgreSQL has 256GB of DDR4 DRAM,
while the server hosting VDMS has 194GB of DDR4 DRAM.
We decided to run the VDMS server in the machine with less DRAM to make
sure MySQL and PostgreSQL had no disadvantage. Previous evaluation
indicated smaller footprint in the case of VDMS when
compared to similar baselines based on MySQL.
To build both MySQL and PostgreSQL on the same machine,
we stored each relational database on separate SSD drives.
Other than the difference in DRAM space and storage needed for two baselines,
the machines are identical.
The client application running the queries and measuring round-trip time
is connected to the server through a 1GB wired link through
a 10GB back-plane switch, same as both servers.
The client application was implemented using Python 3 for both VDMS and the baselines.
Figure~\ref{fig:systems} shows a logical view of the difference between the
interaction of the client application (retrieves metadata and
images) with VDMS (left), the PostgreSQL baseline (middle), and the MySQL baseline (right).

@@

% Another important aspect to note is the storage footprint of the databases.
PostgreSQL and VDMS require comparable amount of storage for metadata which
is more than required by MySQL, shown in Figure~\ref{fig:db_sizes}.
In the case of VDMS, this is space used to store information about
each node/connection.
The Graph Database internal to VDMS (called PMGD) was designed for performance,
especially in environments where persistent memory is present.
This design decision comes as a trade-off for storage footprint, which is
noticeable in our results.
VDMS required 37-41\% more storage than MySQL for storing the same amount
of metadata. On the other hand, PostgreSQL required 1-4\% more storage than VDMS.
The storage footprint may become a factor if storage is a limitation, but
it should also be noted that metadata accounts for less than 2\% of the
overall database size even if we have a 41\% increase in metadata size.
For example, the largest database (both metadata and images) we built
(100M) has around 230GB of metadata and 12TB of images.
In systems where persistent memory is a scarce resource,
the increased storage footprint of PMGD may represent a challenge.
On the other hand, persistent memory is expected to be available
in the order of TBs per server, which should fit the
metadata of intensive use-cases\cite{IntelXPoint15}.

@@

\begin{figure}[ht]
% WAS: begin{figure*}[ht]; textwidth
\includegraphics[width=\columnwidth]{figures/plot_conc_q_1tag_geo_resize_mosaic_results_throughput_db_size}
\caption{Concurrency Analysis on \textit{q2} (\textit{1tag\_geo\_resize}) from our use-case
described in the Experimental Setup Section.
We show each database in different figures for readability reasons.
The figures show the throughput (images per second) when retrieving
resized versions of the images, as the number of concurrent clients increases.
Hardware concurrency (number of physical cores in each system)
is 56.}
\label{fig:concurrency_comparison_q2}
\end{figure}

@@

Another thing to notice is that, as the number of concurrent clients increases,
VDMS throughput continues to increase up to 56 threads, which is
the hardware concurrency of the system.
Also, more parallelism after 56 threads does not increase the delivered throughput
for the larger databases (50M and 100M) but there is a slight improvement in the
other databases (1M, 5M, and 10M).
On the other hand, both baselines seem to deliver less aggregated
throughput after 16 threads except for the 100M case.
In this case, PostgreSQL has a performance spike at 32 clients
and the performance for both baselines is less stable when compared
to VDMS.
Note that most of the throughput for the baseline systems are very close to each other.
This is mostly an effect of the log-scale used, which is needed to clearly
depict the difference between VDMS and the baselines.
Here, the baselines are the full architectures described in
Figure~\ref{fig:systems} (middle and right).

@@

\begin{figure}[ht]
% WAS: begin{figure*}[ht]; textwidth
\includegraphics[width=\columnwidth]{figures/plot_conc_dbsize_10M_mosaic_results_throughput}
\caption{Concurrency Analysis for all the queries using the 10M database).
We show queries in different figures for readability reasons.
Each figure shows the throughput (images per second) when retrieving
resized versions of the images as the number of concurrent clients increases.
Hardware concurrency (number of physical cores in each system)
is 56.}
\label{fig:concurrency_comparison_10M}
\end{figure}

We continue by looking at Figure~\ref{fig:concurrency_comparison_10M}
which shows the images per second delivered by each system
for all queries using the 10M database, as the number of concurrent
clients grows (x-axis). This figure takes a closer look at the
concurrency for each of the six queries.
It is obvious that VDMS consistently outperforms each of the baseline
systems for queries \textit{q1}, \textit{q2}, \textit{q4},
and \textit{q6} as concurrency grows. This figure
shows a similar trend as Figure~\ref{fig:concurrency_comparison_q2}
when it comes to concurrency.  The
VDMS throughput continues to increase up to 56 concurrent clients, which is
the hardware concurrency of the system, for all queries except \textit{q3}.
In this query, the throughput seems to decrease after 32 clients
and both baselines outperform VDMS after 8 concurrent clients.
Prior to 8 clients, VDMS has a slight advantage over PostgreSQL which has
a major performance improvement at 16 concurrent clients. On the other hand,
MySQL maintains its improvement in throughput over PostgreSQL and VDMS over all
concurrent clients. In the case of \textit{q5}, the throughput of the baselines
are less aggregated than those of VDMS up to 16 concurrent clients. When it comes
to low concurrency, the baselines do as good and even better than VDMS with 2 or
4 concurrent clients.  At 16 concurrent client, the performance of the MySQL baseline
begin to stabilize while the throughput of PostgreSQL begins to degrade. However, as
concurrency increases beyond 16 concurrent clients, the difference
in throughput becomes clear, with VDMS reaching its peak performance at
112 concurrent clients.

@@

For \textit{q1} and \textit{q2}, we can appreciate higher performance being delivered by
VDMS when compared to the baselines, and how this improvement is maintained
as the size of the database increase while,
on the other hand, the baseline implementations have a drop in performance
for the 100M database.
For \textit{q3}, we see that VDMS performs best when the database size
is small (1M images), but as the database size increases,
the performance degrades as well.
For this query, MySQL (and in some cases PostgreSQL) outperforms VDMS
for databases larger than 1M.
This also occurs for \textit{q5} but only for databases larger than 10M.
This is entirely attributed to the 2-round process needed for this query,
as we explained before which is more visible in the larger databases.
It is interesting to note that adding filtering by geo-location
(\textit{q5}) decreases the performance as the database sizes scales.
The behaviour is different in the case of \textit{q6},
which also filters by geo-location.
This is attributed to the fact that OR queries involves processing
larger results, and thus does not benefit from the filtering
which is evident when compared to \textit{q4}.

@@

Because of space constraints, we are only showing a subset of our evaluation.
More details about our comprehensive evaluation are available.
\footnote{https://github.com/luisremis/visual\_storm/tree/master/yfcc100m/python/eval/results}
% TODO. Upload the final version of the plots once we are sure we are not making
% more modifications.
